\TheTitle

### 2017-07-14 paper/background.tex
many changes, looks very good

@@

Each repository event contains an ordered list of one or more parents,
except the initial commit which has no parents. The parent relationship
points from the current event toward the initial commit.  Commits have a
single parent, while merges will have at least two parents, the parent
on the current branch followed by the branches being merged into the
current branch. Git allows for the merging of multiple branches
simultaneously in octopus merges, where the order of the parents
represents the order that the branches were specified for being merged.
A \foxtrot\footnote{See
  \url{http://bit-booster.blogspot.ca/2016/02/no-foxtrots-allowed.html}
  for a full description of the issue.} merge occurs when the parent on
the current branch is replaced by a parent from a branch being merged
into the current branch, thus confounding the branch relationships.

The repository events in the DAG are immutable; once a repository event
is created, it can't be changed. Git allows operations to alter commits
and re-order them, but this will create a new commit with a new commit
hash.  This property makes a repository event unable to record the
traversal of merges into the master branch of the repository. A commit
cannot have references in the direction of the merge that merges the
commit, as this reference will change, requiring the commit to be
changed. While this provides some challenge, git eliminates this issue
of the DAG with the \verb|git log --children| command, which traverses
the DAG and inverts the edges to provide the reference from parent to
child.

Under perfect conditions, visualizations of the DAG show how a
repository changes over time; where logical branches are being made,
where merges are happening, and who is making the changes. Most
repositories do not portray the perfect conditions, where there are
relatively few branches and the merges are well-behaved, leaving a
well-defined master branch. Most repositories do not mantain the perfect
structure; either introducing \foxtrot merges, containing many commits,
or many branches, making the DAG difficult to interpret.

>>>>

\section{Background}

@@

Each repository event contains an ordered list of one or more parents,
except the initial commit which has no parents. The parent relationship
points from the current event toward the initial commit.  Commits have a
single parent, while merges will have at least two parents, the parent
on the current branch followed by the branches being merged into the
current branch. Git allows for the merging of multiple branches
simultaneously in octopus merges, where the order of the parents
represents the order that the branches were specified for being merged\footnote{One octopus merge in Linux
  (2cde51fbd0f310c8a2c5f977e665c0ac3945b46d) has 66 parents.}.
A \foxtrot\footnote{See
  \url{http://bit-booster.blogspot.ca/2016/02/no-foxtrots-allowed.html}
  for a full description of the issue.} merge occurs when the parent on
the current branch is replaced by a parent from a branch being merged
into the current branch, thus confounding the branch relationships.\dmg{this would benefit from a diagram, since
most people will have no idea what you are talking about.}

The repository events in the DAG are immutable; once a repository event
is created, it can't be changed. \dmg{the next sentences should be repository events, not commits}
Git allows operations to alter commits
and re-order them, but this will create a new commit with a new commit
hash.  This property makes a repository event unable to record the
traversal of merges into the master branch of the repository. A commit
cannot have references in the direction of the merge that merges the
commit, as this reference will change, requiring the commit to be
changed. \dmg{BE CONSISTENT, Git or git?} git provides
the command \verb|git log --children| which traverses
the DAG and prints the inverted edges relationship between events.

Under perfect conditions, visualizations of the DAG should show how a
repository changes over time; where logical branches are being made,
where merges are happening, and who is making the changes. Most
repositories are relatively simple, with
relatively few, small, branches with a clear, well defined
master branch. However, in large, active repositories it becomes very difficult for developers to use the DAG
to identify
the master branch, and as a consequence, where branches start and are merged back in their
way to master.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "lineval.tex"
%%% End:

### 2017-07-14 paper/introduction.tex
many changes, looks very good

@@

The DAG is able to provide a meaningful visualization in smaller
projects; it enables users to see when changes are made, when these
changes are merged, how each branch is interacting, and the point where
a branch forks from the master branch. In large modular projects, like
the Linux kernel, the DAG becomes a mess of merges and commits
(Figure~\ref{fig:gitk}) losing its visual meaning. In some cases, the
Linux kernel is simply too large for the system to generate a
visualization; Github provides a DAG view for many projects, but is
unable to display the visualization for projects at the scale of the
Linux of the kernel (Figure~\ref{fig:gitfail}). Between 60k and 70k new
commits are created for the Linux project every year; according to
previous work\cite{German2015}, a commit takes a median of 30 days from
the time it is authored until it arrives in the official repository. The
snapshot of the kernel tomorrow may be different than the snapshot from
today, containing new commits authored in the past; distinguishing these
new commits from the commits in the snapshot from today is not trivial.

@@

This paper makes three contributions; first, we describe a method of
converting the DAG of the Linux repository into the \mt of the
repository, that represents the path used by a commit to reach the
master branch; second, we present an implementation enabling the
inspection and visualization of merges in the Linux project using the
\mt modle; finally, we validate the \mt model and the implementation of
the visulization through a controlled user study. We further discuss the
issues of generalizing the model to other repositories, but present an
updated version that improves the performance of the algorithm by
pruning the DAG.\@

>>>>

\section{Introduction}

@@

The DAG is able to provide a meaningful visualization in smaller
projects; it enables users to see when changes are made, when these
changes are merged, how each branch is interacting, and the point where
a branch forks from the master branch. In large modular projects, like
the Linux kernel, the DAG becomes a mess of merges and commits
(Figure~\ref{fig:gitk}) losing its visual meaning. In some cases, the
Linux kernel is simply too large for the system to generate a
visualization; Github provides a DAG view for a repository, but is
unable to display the visualization for projects at the scale of the
Linux of the kernel (Figure~\ref{fig:gitfail}). Between 60k and 70k new
commits are created for the Linux project every year; according to
previous work\cite{German2015}, a commit takes a median of 30 days from
the time it is authored until it arrives in the official repository. The
snapshot of the kernel tomorrow may be different than the snapshot from
today, containing new commits authored in the past; distinguishing these
new commits from the commits in the snapshot from today is not trivial.

@@

\dmg{add to the abstract an abbreviation of this sentence: see my comment in the abstract }
This paper makes three contributions; first, we describe a method of
converting the DAG of the Linux repository into the \mt of the
repository, that represents the path used by a commit to reach the
master branch; second, we present an implementation enabling the
inspection and visualization of merges in the Linux project using the
\mt modle; finally, we validate the \mt model and the implementation of
the visulization through a controlled user study. We further discuss the
issues of generalizing the model to other repositories, but present an
updated version that improves the performance of the algorithm by
pruning the DAG.\@

### 2017-07-14 paper/lineval.tex
Adding background section
many changes, looks very good

\section{Introduction}

@@

\section{Merge-Tree model}
\label{sec:mergetree}

@@

In our visualizations, we leverage the merge-tree model described in
Section~\ref{sec:mergetree}. In this model each commit is either already
in the master branch, or is part of a tree which is rooted in the merge
that merged it into the master branch.  Each commit, whether a merge or
non-merge, has only one successor; the root of each tree has none as it
was made by Linus Torvalds directly into the master branch. Non-merge
commits contain the metadata for the changes made.  This metadata
includes the files changed, the lines added and removed from each file,
the author, the date the commit was merged into the merge that led to
being merged into the kernel, the date the commit was authored, the
patch, and the commit log. Merges contain less metadata, only storing
the author of the merge, the log, the commit date, and the author date,
and potentially, changes necessary to address conflicts during the
merge. The details of the model are outlined in \cite{German2015}.

@@

\section{Related Works}
\label{sec:related_works}

@@

\section{Controlled User Study}
\label{sec:study}

>>>>

@@

\label{sec:background}

\input{background.tex}

@@

In our visualizations, we leverage the merge-tree model described in
Section~\ref{sec:mergetree}. In this model each commit is either already
in the master branch, or is part of a tree which is rooted in the merge
that merged it into the master branch.  Each commit, whether a merge or
non-merge, has only one successor; the root of each tree has none as it
was made by Linus Torvalds directly into the master branch. Non-merge
commits contain the metadata for the changes made.  This metadata
includes the files changed, the lines added and removed from each file,
the author, the date the commit was merged into the merge that led to
being merged into the kernel, the date the commit was authored, the
patch, and the commit log. Merges contain similar metadata,
the author of the merge, the committer of the merge, if it has been rebased, the log, the commit date, and the author date,
and potentially, changes necessary to address conflicts during the
merge. The details of the model are outlined in \cite{German2015}.

@@

@@

### 2017-07-14 paper/model.tex
many changes, looks very good

% vim:set et sw=2 ts=4 tw=72:
The \mt model inverts the DAG, so that reference point from the parent
to the child instead of from child to parent. This provides a clear path
from a commit to the closest merge into the master branch. We define the
distance as a measure of time from when the commit was initially
commited to the date it was merged into the project. The merge into the
master branch becomes the root of the tree, and the commits are the
leaves. Any merges that that a commit must pass through to reach the
master branch become an inner node of the tree. The structure is able to
help a user identify the path that the commit took to reach the master
branch, as well as understand how commits were grouped together in order
to be integrated. In this model, not only has the DAG been inverted, but
the entire notion of parent-child relationship has been reversed. The
terminology surrounding the parent-child relationship will prove to be
confusing through the rest of the paper, when we are referring to the
\mt, the parent is the next node toward the merge into the master
branch, or the root of the tree. When referring to the DAG, the parent
relationship is in the opposite direction, from the root toward the
branch-point. In addition to identifying the path that a commit took to
being merged, we are able to aggregate the commit metadata at merges as
we retain the parent-child and child-parent relationship for each event.

A short example: assume the commits represented in
Figure~\ref{fig:repoEvents} show the sequence of events in a repository.
The sequence starts with the initiall commit in the master branch of the
master repository at time $t_0$. Repository event 1 is a commit, which
gets forked into a separate repository, \textit{Repo A}, where another
commit is made, event 2. Event 5 is a merge event, merging event 2, 3,
and 4 into \textit{Repo A}. Event 5 is branched from, event 6 happens in
the new branch, while event 7 is added simultaneously to the original
branch in \textit{Repo A}. Events 11 and 12 are both merge events,
merging changes made in \textit{Repo A} into the master branch of the
master repository.

@@

@@

>>>>

% vim:set et sw=2 ts=4 tw=72:

\section{Merge-Tree model}

\label{sec:mergetree}

The \mt model abstracts the DAG of the repository into a set of \mt{}s. Each \mt is rooted at a merge into the master
branch. The children of this tree are the commits and merges that are merged together into master by this merge. A \mt is build
recursively with every merge in the \mt merging a sub-\mt. Any merges that a commit must pass through to reach the
master branch become an inner node of the tree. 
Effectively, a \mt is a tree that
shows the path that commits follow in their way to the master branch of a repository.
The \mt{}s also help understand how commits were grouped together in order
to be integrated. In this model, not only has the DAG been inverted (and simplified), but
the entire notion of parent-child relationship has been reversed.

Please note that when we are referring to the
\mt, the parent is the next node toward the merge into the master
branch, or the root of the tree. When referring to the DAG, the parent
relationship is in the opposite direction, from the root toward the
branch-point. In addition to identifying the path that a commit took to
being merged, we are able to aggregate the commit metadata at merges as
we retain the parent-child and child-parent relationship for each event.

To illustrate this model we will use a small example:
assume the commits represented in
Figure~\ref{fig:repoEvents} show the sequence of events in a repository.
The sequence starts with the initiall commit in the master branch of the
master repository at time $t_0$. Repository event 1 is a commit, which
gets forked into a separate repository, \textit{Repo A}, where another
commit is made, event 2. Event 5 is a merge event, merging event 2, 3,
and 4 into \textit{Repo A}. Event 5 is branched from, event 6 happens in
the new branch, while event 7 is added simultaneously to the original
branch in \textit{Repo A}. Events 11 and 12 are both merge events,
merging changes made in \textit{Repo A} into the master branch of the
master repository.

@@

\dmg{I think you abbreviated this section too much. It will be helpful to describe the algorithm
at a very high level,as in the vissoft paper.}

@@

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "lineval.tex"
%%% End:

### 2017-07-14 paper/relatedwork.tex
many changes, looks very good

version control systems monitor the development lifetime of software
projects. This makes the version control system vital in providing
information about how a software project is being developed. To do this,
the user must be able to gain a clear understanding of how commits are
being integrated into the project, along with understanding what changes
these commits are bringing with them. These are the tasks that we set
out to solve with the \mt model. Ideally, we could compare our model
against another that is designed with these goals. To our knowledge, we
know of no git repository visualzation tool that has these specific
goals. This may be due to issues with finding the master branch of the
repository which is a non-trivial task, or due to other factors.

@@

% TODO: Check this -- Look into source tree
Not included in an academic project are both the git commandline with
the \verb|git log| command, and Gitk, which is a graphical tool shipped
with git. Many other tools are designed with a similar visual metaphore
as what is presented in Gitk, including Git kraken, Github Desktop and
web interfaces, the Git Lab web intereface, the Bitbucket web interface,
and many others. The Gitk interface is built around the central DAG
viewer. The DAG displays the repository events on their respective
branches, the author, and the date that the event was authored. A user
can select an event to view addition information about it, including the
full commit log, the parents and children of the event, and the diff
generated if modifications were made. In the case of the Linux kernel,
merges are not resolved as merge conflicts in the master repository, but
by the developers prior to merging. With this development model, no
merges will have modified any files, and will not show a diff. While the
merge does not make changes directly, the effects of merging the
commits persist, but Gitk is unable to provide an aggregated view of the
files modified, the patches for those files, and the authors who
contributed commits to that merge.

>>>>

\section{Related Works}
\label{sec:related_works}

Version control systems monitor the development lifetime of software
projects. This makes the version control system vital in providing
information about how a software project is being developed. To do this,
the user must be able to gain a clear understanding of how commits are
being integrated into the project, along with understanding what changes
these commits are bringing with them. These are the tasks that we set
out to solve with the \mt model. Ideally, we could compare our model
against another that is designed with these goals. To our knowledge, we
know of no git repository visualzation tool that has these specific
goals. This may be due to issues with finding the master branch of the
repository which is a non-trivial task, or due to other factors.

@@

% TODO: Check this -- Look into source tree
Git itself is shipped with summarization and visualization tools:
the \verb|git log| command and Gitk, which is a graphical tool for the purpose of browsing
the DAG of the repository.
Many other tools are designed with a similar visual metaphore
as what is presented in Gitk, including Git kraken, Github Desktop and
web interfaces, the Git Lab web intereface, the Bitbucket web interface,
and many others. The Gitk interface is built around the central DAG
viewer. The DAG displays the repository events on their respective
branches, the author, and the date that the event was authored. A user
can select an event to view addition information about it, including the
full commit log, the parents and children of the event, and the diff
generated if modifications were made. In the case of the Linux kernel,
merges are not resolved as merge conflicts in the master repository, but
by the developers prior to merging. With this development model, no
merges will have modified any files, and will not show a diff. While the
merge does not make changes directly, the effects of merging the
commits persist, but Gitk is unable to provide an aggregated view of the
files modified, the patches for those files, and the authors who
contributed commits to that merge.

### 2017-07-14 paper/study.tex
Cleaning up todos
Removing some TODO's, removing some hyphenations.
many changes, looks very good

The purpose of the study is to quantitatively and qualitatively evaluate
the effectiveness of \tool compared to the DAG-based visualizations
found in Gitk or the git command line. The study is broken into three
task-based sections; conceptual tasks, summarization tasks, and
participant opinions. The study was performed in June of 2017 in
Kingston ON, in a controlled environment running Ubuntu 14.04.
Participants were allowed to use both Gitk and the git command line
tools together, which we will refer to as the git tools or as Gitk. Then
they were asked to perform the same tasks using \tool.

\evan{Do I need to say where the study took place? I see it showing up
  in other papers, but it makes it kind of obvious about the lab I'm
  talking about, doesn't it?}

The primary goal of the study is to determine if \tool is more capable
of providing the participants of our study with a better conceptual
understanding of how a commit is integrated into the kernel, and a
better summarization of various metrics about the files and authors
involved with the integration. The conceptual questions are to determine
if the DAG visualization provides enough information to determine how a
commit, and any related commits, are merged into a project. \tool
displays this information directly in the form of a tree, so we will not
be performing a comparison on this section, only evaluating the results
from Gitk. The summarization tasks compare \tool and Gitk, with the
participants using one tool followed by the other on two commits from
merge-trees of differeing sizes. In both of these sections, the
participant is given a task that they are to find an answer to; we
evaluate the time it took, if the answer was correct, and how far from
the right answer they were. The user opinions provide us with a clear
comparison between the tools from the point-of-view of the participant,
providing us information about which aspects of each tool they prefered.

% TODO: Add the participant profile

@@

@@

\textbf{The summarization tasks}, listed in Table~\ref{tab:summarization_tasks},
are designed to compare the ability of the participants to summarize
information about the two commits used in the study when using git tools
verses \tool. The tasks are broken into four sets based on what is being
summarized: merges, authorship, files, and modules. The order that the
participants perform each task is randomized within each task set, and
the order of the task sets are then randomized with the exception of the
merges task set which will always come first. We measure and evaluate
three metrics for each task: correctness, accuracy, and timing.
Correctness measures whether the answer was correct or incorrect.
Accuracy measures how far from being correct the provided answer was.
Timing is the duration, in seconds, that the participant took to answer
the task. We begin timing the task at the end of the question, and stop
timing before the last modification to the answer. For this reason, it
is possible for times to overlap between tasks if the participant
changes their answer after answering another question. This will also
modify the answer used for measuring the accuracy. We measure the timing
from the screen capture.

\evan{Is this better, or do you still want me to move the information
  about the randomization to the bottom?}

@@

We include two commits in the study to detect differences between merge
tree sizes. The order that the commits are presented to each participant
is randomized. We analyzed the 15096 merge-trees from April 16th 2005 to
October 14th 2014, which corresponds to the releases Linux 2.6.12-rc3
and Linux 3.17-rc1. 25\% of the trees contain at most a single commit
not including the root, while 50\% of the trees contain at most seven
nodes non-root nodes.
% TODO: get the number for this if there is time
75\% of the trees contain at least 51 nodes, and the largest tree
contains 7217 nodes.

@@

Five participants had only worked with personal and academic
repositories. Two participants had only worked with school and
professional repositories, and two had only worked with professional
repositories. One participant worked with personal, academic, and
professional repositories, one participant had only worked with personal
repositories, and one participant had worked with personal, academic,
and school repositories.

\evan{Do you want more information about the participants?}

@@

@@

Summaries for correctness, accuracy, and timing are shown in
tables~\ref{tab:correctness_table}.

% TODO: add results for other metrics

The full results are as follows;

>>>>

\section{Controlled User Study}
\label{sec:study}

The purpose of the study is to quantitatively and qualitatively evaluate
the effectiveness of \tool compared to the DAG-based visualizations
found in Gitk or the git command line. The study is broken into three
task-based sections; conceptual tasks, summarization tasks, and
participant opinions. The study was performed during June of 2017
%in Kingston ON,
in a controlled environment running Ubuntu 14.04.
Participants were allowed to use both Gitk and the git command line
tools together, which we will refer to as the git tools or as Gitk. Then
they were asked to perform the same tasks using \tool.

\dmg{I thought you had randomized the order of tool use? also clarify that they were done with different commits to
  avoid bias}

% \evan{Do I need to say where the study took place? I see it showing up
%   in other papers, but it makes it kind of obvious about the lab I'm
%   talking about, doesn't it?}

The primary goal of the study is to determine if \tool is more capable
of providing the participants of our study with a better conceptual
understanding of how a commit is integrated into the kernel, and a
better summarization of various metrics about the files and authors
involved with the integration. The conceptual questions are to determine
if the DAG-based visualizations (gitk and git command line) provide enough information to determine how a
commit, and any related commits, are merged into a project. \tool
displays this information directly in the form of a tree, so we will not
be performing a comparison on this section, only evaluating the results
from Gitk. The summarization tasks compare \tool and Gitk, with the
participants using one tool followed by the other on two commits from
merge-trees of differeing sizes. In both of these sections, the
participant is given a task that they are to find an answer to; we
evaluate the time it took, if the answer was correct, and how far from
the right answer they were. The user opinions provide us with a clear
comparison between the tools from the point-of-view of the participant,
providing us information about which aspects of each tool they prefered.

@@

\dmg{I don't like how you talk about the DAG comprehension. Few paragraphs above I replaced it
  with the DAG-based tools, which might be a better description, since you are evaluating tool
vs tool}

@@

\textbf{The summarization tasks}, listed in Table~\ref{tab:summarization_tasks},
are designed to compare the ability of the participants to summarize
information about the two commits used in the study when using git tools\dmg{be consistent, either you say dag-based
  tools, or git tools, either one is fine, but be consistent}
versus \tool. The tasks are broken into four sets based on what is being
summarized: merges, authorship, files, and modules. The order that the
participants perform each task is randomized within each task set, and
the order of the task sets are then randomized with the exception of the
merges task set which will always come first. We measure and evaluate
three metrics for each task: correctness, accuracy, and timing.
Correctness measures whether the answer was correct or incorrect.
Accuracy measures how far from being correct the provided answer was.
Timing is the duration, in seconds, that the participant took to answer
the task. We begin timing the task at the end of the question, and stop
timing before the last modification to the answer. For this reason, it
is possible for times to overlap between tasks if the participant
changes their answer after answering another question. This will also
modify the answer used for measuring the accuracy. We measure the timing
from the screen capture.

\evan{Is this better, or do you still want me to move the information
  about the randomization to the bottom?}
\dmg{I didn't see anything about choosing what commits to inspect here... but this section looks good}

@@

We include two commits in the study to detect differences between merge
tree sizes. The order that the commits are presented to each participant
is randomized. We analyzed the 15096 merge-trees from April 16th 2005 to
October 14th 2014, which corresponds to the releases Linux 2.6.12-rc3
and Linux 3.17-rc1. 25\% of the trees contain at most a single commit
not including the root, while 50\% of the trees contain at most seven
nodes non-root nodes. 75\% of the trees contain at least 51 nodes, and
the largest tree contains 7217 nodes.

@@

Five participants had only worked with personal and academic
repositories. Two participants had worked with school and
professional repositories, and two had only worked with professional
repositories. One participant worked with personal, academic, and
professional repositories, one participant had only worked with personal
repositories, and one participant had worked with personal, academic,
and school repositories.
\dmg{you made it confusing: just say: how many had experience with
  professional repositories and remove the above
}
\evan{Do you want more information about the participants?}

@@

\dmg{BIG QUESTION}
for each task, did they do always git tools then linvis for the same commit?  if so, there is potential for learning
bias: they learn the structure of the commits with git tools, so therefore it was easier for them to do
them with linvis.
\dmg{END QUESTION}

\dmg{I'll stop here today}

@@

Summaries for correctness, accuracy, and timing are shown in
tables~\ref{tab:summarization_table}.
The full results are as follows;
